---
title: '[R] Crop model with tidyverse (Part III)'
author: 'Benjamin Nowak'
date: '2021-07-14'
slug: r-crop-model-with-tidyverse-part-iii
categories: ["R"]
tags:
  - model
  - optimization
  - tidyverse
---


```{r setup, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# Set Working directory for Notebook
knitr::opts_knit$set(root.dir='/Users/bnowak/Documents/231220/WD/AnalyseSyst/')
```


```{r echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
# Load data 
input <- read.table(file="Weather_DesMoines.csv", header=TRUE, sep=";",dec=".")
# Clean data
data<-input%>%
  dplyr::select(
    tas = T_DAILY_MEAN,
    rsds = SOLARAD_DAILY  
  )%>% 
  dplyr::mutate(                
    day_number = row_number()     
  )%>%
  dplyr::filter(
    day_number>=92                # Keep only data after April 1st (included)
  )

# Outside the function:
# Rquired parameters to compute C
# Light extinction coefficient
K <- 0.56
# Individual leaf area (m-2)
S <- 0.05
# Plant density (m-2)
d <- 90000/10000

# Model function
model_fun <- function(
  name,           # Scenario name 
  data,           # Climatic variables to be used as inputs
  GDD_1leaf,      # Thermal requirement for the emergence of one leaf
  C,              # C=f(K,S,d)
  RUE,            # Radiation use efficiency (gDM.MJ-1)
  nthresh         # Number of leaves before grain filling
  ){      
  # Set parameters (without GDD_1leaf)
  max_nleaf<-20
  T0<-6 
  f<-0.5      # Active fraction of incoming radiation
  frac<-0.7   # Fraction of Net Primary Productivity dedicated to grain
  
  # Estimate yield
  model<-data%>%
    dplyr::mutate(
      TT=dplyr::case_when(
        tas<T0~0,
        tas>=T0~tas-T0
    ))%>%  
    mutate(
      GDD = cumsum(TT)
    )%>%
    mutate(
      pot_nleaf = GDD/GDD_1leaf
    )%>%
    mutate(
      nleaf = case_when(
        pot_nleaf<=max_nleaf~round(pot_nleaf),
        pot_nleaf>max_nleaf~max_nleaf
      )
    )%>%
     # Incoming photosynthetic active radiation (MJ.m-2.day-1)
    mutate(
      PAR_inc = f*rsds
    )%>%
    # Absorbed PAR by the canopy (MJ.m-2.day-1)
    mutate(
      APAR = PAR_inc*(1-exp(-C*nleaf))
    )%>%
    # Net primary productivity dedicated to the aboveground biomass 
    mutate(
      NPP = RUE*APAR
    )%>%
    # Sum of aboveground biomass
    mutate(
      biom = cumsum(NPP)
    )%>%
    # Net primary productivity dedicated to the variable grain
    mutate(
      NPPgrain = case_when(
        nleaf<nthresh ~ 0,
        nleaf>=nthresh ~ frac*NPP
      )
    )%>%
    # Total grain production (g.m-2)
    mutate(
      grain = cumsum(NPPgrain)
    )%>%
    # Total grain production (t.ha-1)
    mutate(
      grain_t = grain/100
    )%>%
    add_column(                                # To add scenario name to data
      Scenario = name                          # (set 'name' in argument)
    )
  return(model)
}

```

**Goals:**   

  * Calculate Root Mean Square Error
  * Optmize model 
  * Plot results

One common measure of the differences between values predicted by a model and the values observed is the Root Mean Square Error. In our case, it is calculated as follows:

  $$ \quad\text{(Eq. 1)}\quad  RMSE = \sqrt\frac{\sum_{i=1}^N(Y_{i}^{Our Model}-Y_{i}^{Ref Model})^{2}}{N}$$

With:<br> 
$Y_{i}^{OurModel}$: final yield estimated by our model ($t.ha^{-1}$)<br>
$Y_{i}^{RefModel}$: final yield estimated by reference model ($t.ha^{-1}$)<br>
$N$: total number of simulations<br>


# Acknowledgement

Many thanks to Bruno Ringeval for taking the time to answer my questions.

# Reference
* Ringeval, B. *et al.* (**2021**). Potential yield simulated by global gridded crop models: using a process-based emulator to explain their differences. Geosci. Model Dev., 14, 1639–1656. https://doi.org/10.5194/gmd-14-1639-2021